%!TEX root = ../document.tex
\chapter*{Acknowledgements}
\thispagestyle{empty}

The work here presented is delivered as a final thesis report at Lisbon Tech, University of Lisbon, Portugal. This thesis is part of a 2 years master programme with focus on peer-to-peer networks, cloud computing and mobile applications.

I want to show my appreciation for all the support given from my family, specially to my father, who has worked hard his entire life, even with all the challenges that were presented, to enable me and my sister to go through college. He taught me that preseverance, honesty and humility trumps every challenge, that we never give up.

To my friends, Camille Coyac, Nuno Job, Adam Brault, Diogo Mónica, Pedro Teixeira,  Thiago Dantas, Ben Acker, Liliana Ângelo and André Albuquerque, I deeply thank you for being such an inspiring force through my life, for believing and showing interest, providing guidance and for being present in times of need.

A special thank you to my professor and thesis coordinator, Luís Veiga, who has proven to be one of the most interesting humans I've met and with always amazing, rapid and open minded interest and feedback to the work we developed during the master thesis.

I also want to show my appreciation to all the people that appeared during this journey, that through their presence or an informal conversation, have contributed to the success of this work, these are: Feross Aboukhadijeh, Adam Baldwin, Ilya Gregorik, Phillip Hancke, Igor Soarez, Bernardo Simões, Philip Roberts and Nathan Fritz.

\vspace{15pt}
\vfill
\begin{flushright}
    \begin{minipage}{8cm}
        \begin{center}
            10th of May, 2015, Lisbon \\ 
            David Dias
        \end{center}
    \end{minipage}
\end{flushright}

\cleardoublepage

\chapter*{}
\thispagestyle{empty}
\vfill
\vfill
\begin{flushright}
    \begin{minipage}{8cm}
        \begin{center}
            {\Large\bf --To all of the first followers, you undoubtly changed my life.}

        \end{center}
    \end{minipage}
\end{flushright}

\cleardoublepage
